---
title: "RFM score as input for K-means clustering"
author: "Minh Trung DANG"
date: "08/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr)
```

```{r}
data <- read_csv("customerFMCG.csv")
```
```{r}
str(data)
```

Data desciption

Invoice_No (Numeric): Invoice no for each transaction
Stock_Code (Numeric): Unique stock code for the items
Product_Category(Categorical):Product category details
Invoice_Date(Numeric):Date on which invoice was generated
Customer_ID (Numeric):Unique customer id 
Amount (Numeric): Invoice Amount
Country(Categorical):Country detail
l_date(Numeric):Last date of invoice in 2011


```{r}
library(DataExplorer)
```

```{r}
plot_missing(data)
```

```{r}
head(data,10)
```

```{r}
str(data)
```

```{r}
library(dplyr)
```

Counting unique 
```{r}
n_distinct(data$Invoice_No) 
```

```{r}
n_distinct(data$Stock_Code) 
```

```{r}
data %>%
  group_by(Product_Category) %>% 
  count()
```

```{r}
library(pander)
library(ggplot2)
library(scales)
```

```{r}
df_category <- data %>% 
  group_by(Product_Category) %>% 
  summarise(Count =  n()) %>% 
  arrange(-Count) %>% 
  mutate(Percentage = round(Count*100/sum(Count),2),
         label =  percent(Percentage/100))
  

df_category
```


```{r, fig.height= 5, fig.width = 6.5}
ggplot(df_category,
       aes(x = reorder(Product_Category, -Percentage), 
           y = Percentage)) +
  geom_bar(stat = "identity",
           fill = "royalblue4",
           width = 0.80) +
  geom_text(aes(label = paste0(Percentage,"%")), 
                               vjust = -0.3,
                               size = 4,
                               fontface = "bold") +
  labs(x = "Category",
       y = "Distribution of product category (%)") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12, face = "bold"),
        plot.caption = element_text(color = "grey44",size = 12,face = "italic"),
        legend.text = element_text(colour="black", size = 12))
```

```{r}
n_distinct(data$Customer_ID)
```

There are 3813 unique custmomers. 15335 unique invoice numbers. 


```{r}
head(data$l_Date)
```

```{r}
data2 <- data %>% 
  mutate(Invoice_Date =  as.Date(Invoice_Date, "%m/%d/%Y"))
```

```{r}
str(data2)
```

```{r}
data2 %>% 
  group_by(Country) %>% 
  count()
```
All these records belong to only one country. We will not neet this column.


```{r}
summary(data2$Amount)
```


```{r}
hist(data2$Amount)
```

```{r}
 ggplot(data2, aes(x="", y=Amount)) + 
  geom_boxplot()
```

There are outliers. 

```{r}
data2 %>% 
  group_by(l_Date) %>% 
  count()
```
This columne is redandunt. 

Building RFM model

```{r}
R_table <- aggregate(Invoice_Date ~ Customer_ID, data2, FUN = max)
```

```{r}
data2 %>% 
  group_by(Customer_ID) %>% 
  summarise(max = max(Invoice_Date))
```

```{r}
head(R_table)
```

```{r}
NOW <- as.Date("2011-12-12", "%Y-%m-%d")
NOW
```


```{r}
R_table$R <- as.numeric(NOW - R_table$Invoice_Date)
```

```{r}
head(R_table)
```


```{r}
RFM_data <- data2 %>% 
  group_by(Customer_ID) %>% 
  summarise(Recency = as.numeric(NOW - max(Invoice_Date)),
            Frequency = length(Invoice_Date),
            Monetary = sum(Amount)) 
```

```{r}
str(RFM_data)
```


```{r}
head(RFM_data)
```

RFM scoring

Rsegment 1 is very recent while Rsegment 5 is least recent

In this step, scoring the RFM data is done by using the quantile method. The scoring is in a range of 1 to 5. 
In a R segementm 1 is very recent whilke 5 is the least recent score.
Fsegment 1 is least frequent while Fsegment 5 is most frequent

```{r}
RFM_data <- RFM_data %>% 
  mutate(Rsegment = findInterval(Recency, quantile(Recency, c(0.0, 0.25, 0.50, 0.75, 1.0))),
         Fsegment = findInterval(Frequency, quantile(Frequency, c(0.0, 0.25, 0.50, 0.75, 1.0))),
         Msegment = findInterval(Monetary, quantile(Monetary, c(0.0, 0.25, 0.50, 0.75, 1.0))),
         R_F_M = paste(Rsegment, Fsegment, Msegment),
         Total_RFM_Score = c(Rsegment +  Fsegment + Msegment))

```

```{r}
head(RFM_data)
```

To keep only selected variables from RFM data

```{r}
clus_df2 <- RFM_data[,c(5,6,7)]
```



Applying K-means

```{r}
set.seed(123)
km2 <- kmeans(clus_df2, centers = 5, nstart = 30)
```


```{r}
str(km2)	
```
```{r}
library(factoextra)
```

```{r, fig.height= 7.5, fig.width= 7.5}
fviz_cluster(km2, data = clus_df2)
```


```{r}
finalclus2 <- cbind(clus_df2,km2$cluster)
```

```{r}
finalclus2 %>% 
  group_by(km2$cluster) %>% 
  count()
```


